Guarded Recursion in Agda via Sized Types

### 2019-01-25 code-oneclock-paper/latex/Introduction.tex
added a couple of comments
comments on Clouston et al. and Abel & Vezzosi in intro

Utilizing techniques for representing ``type theory in type theory''
\cite{Chapman09,AltenkirchK16}, we present
an Agda formalization of the syntax of a simple type theory with
guarded recursion. This object language, that we call \GTT, is a variant of Atkey and
McBride's type system for productive coprogramming
\cite{atkey2013productive} in which the clock context can contain at
most one clock. The types of \GTT\ include the aformentioned \IC{▻} modality and a
\IC{□}  modality, the nameless analogue of  Atkey and
McBride's universal clock quantification.

Afterwards we develop a categorical semantics
for \GTT\ using sized types. More precisely, we define a presheaf
model where the type formers of simply typed lambda calculus are
evaluated using the standard Kripke semantics. Typically the semantics
of type theories for guarded recursion is given in the topos of
trees or variations thereof \cite{BMSS-synthetic,Mogelberg14,MannaaM18}.
Here, in order to make clear the connections between guarded recursion and
sized types, we take the preorder of sizes as the indexing
category of our presheaves. This means that types and contexts of \GTT\ are
interpreted as antitone sized types. 
The well-order relation on sizes is fundamental for giving a
terminating definition of the semantic fixpoint combinator.

@@

This paper only include the essential parts of our Agda
formalization. The full code is available at
\url{https://github.com/niccoloveltri/agda-clott}. The formalization
uses Agda 2.5.4.1 and Agda standard library 0.16.

>>>>

Utilizing techniques for representing ``type theory in type theory''
\cite{Chapman09,AltenkirchK16}, we present an Agda formalization of
the syntax of a simple type theory with guarded recursion. This object
language, that we call \GTT, is a variant of Atkey and McBride's type
system for productive coprogramming \cite{atkey2013productive} in
which the clock context can contain at most one clock. The types of
\GTT\ include the aformentioned \IC{▻} modality and a \IC{□} modality,
a nameless analogue of Atkey and McBride's universal clock
quantification. Clouston et al. \cite{CloustonBGB15} have studied a
guarded variant of lambda calculus extended with a \IC{□} operation,
which they call ``constant'' .  Our object calculus differs from
theirs in that our judgments are indexed by a clock context, which can
be empty or containing exactly one clock. In this respect \GTT\ is
closer in spirit to Atkey and McBride's language. The design of \GTT\
also has the benefit of allowing a more pleasant introduction rule for
\IC{□} than Clouston et al.'s.

Afterwards we develop a categorical semantics
for \GTT\ using sized types. More precisely, we define a presheaf
model where the type formers of simply typed lambda calculus are
evaluated using the standard Kripke semantics. Typically the semantics
of type theories for guarded recursion is given in the topos of
trees or variations thereof \cite{BMSS-synthetic,Mogelberg14,MannaaM18}.
Here, in order to make clear the connections between guarded recursion and
sized types, we take the preorder of sizes as the indexing
category of our presheaves. This means that types and contexts of \GTT\ are
interpreted as antitone sized types. 
The well-order relation on sizes is fundamental for constructing a
terminating definition of the semantic fixpoint combinator.

Abel and Vezzosi \cite{AbelV14} have also formalized a simple type theory
extended with the later modality in Agda. They focus on operational
properties of the calculus and give a certified proof of strong
normalization. In our work we consider a richer type system and we
formalize its denotational semantics in Agda. Nevertheless, we do not
investigate the dynamic behaviour of our calculus, which we leave to
future work.

@@

This paper only include the essential parts of our Agda
formalization. The full code is available at
\url{https://github.com/niccoloveltri/agda-clott}. The formalization
uses Agda 2.5.4.1 and Agda standard library 0.16. The paper is
extracted from a literate Agda file, which implies that all the
displayed code has passed Agda's type-check.
